Comparing Different Planned Missingness Designs in Longitudinal Studies

### 2018-06-08 Documentation/manuscript.tex
Had to commit

% Predefine some figures

@@

@@

\author{Josephine Wood \and Gregory J. Matthews \and Nicholas Illenberger \and Jennifer Pellowski \and Ofer Harel}

\institute{J. Wood \and G. J. Matthews
          \at Loyola University Chicago \\
          1032 W. Sheridan Rd.\\
          Chicago IL, 60660 USA\\
          Tel: (773)508-3558
          \and
          N. Illenberger \and O. Harel 
				\at University of Connecticut \\
						215 Glennbrook Rd. Unit 420\\
						Storrs CT, 06269-4120 USA
              Tel: (860)486-6989 \\
		      \email{ofer.harel@uconn.edu}  
					\and
					J. Pellowski
					\at International Health Institute\\ 
					Brown School of Public Health 
					\and J. Pellowski 
					\at Department of Psychiatry and Human Behavior\\
					The Warren Alpert Medical School of Brown University.  
}

@@

\section{Introduction}
\label{intro}
Planned Missingness (PM) designs are a type of experimental design in which some of the data is predetermined to be missing. While it may seem counterproductive to deliberately omit data from the collection process, PM designs have been shown to reduce both participant burden \citep{graham2006planned, saris20048} and the costs of a study \citep{graham2001planned}. Moreover, PM designs can actually increase the quality of collected data by increasing the breadth of a study \citep{little2013planned} as well as decreasing the effects of assessment reactivity and rates of unplanned missingness \cite{harel2015designed}. We further note that while PM designs decrease the efficiency of estimators \citep{rhemtulla2016asymptotic}, decreasing the cost per participant may allow researchers to enroll more participants in a study, offsetting some of the efficiency loss. \par

There have been many different incarnations of the PM design, each crafted to best suit the study at hand. For example, PM designs have been used in the study of longitudinal growth curves \citep{graham2001planned, mcardle1997expanding}, business research \citep{shoemaker1973principles}, behavioral studies \citep{harel2015designed}, biostatistics \citep{andres2006partial, wacholder1994partial}, and educational assessments \citep{zeger1997efficient, sirotnik1977incidence}. This paper focuses on two of the simplest and most applicable PM designs; the Split Form design \citep{raghunathan1995split} and the Wave Missingness design \citep{little2013planned}. While this paper covers these designs in more detail later, broadly speaking, the Split Form design entails measuring fewer points on a single measurement occasion \citep{raghunathan1995split}. The Wave Missingness design, on the other hand, requires missing an entire measurement occasion within a longitudinal study \citep{graham2001planned}.

@@

Split Form missingness designs \citep{raghunathan1995split}, also  known as Multi-Form \citep{little2013planned} or Matrix Sampling designs \citep{thomas2006evaluation}, are very aptly named. Instead of giving participants one large survey, researchers create multiple forms, each containing a subset of the full survey's questions. How these subsets are chosen varies. A common form setup is three-form Block Design \citep{graham1996maximizing} in which the full survey is split into four blocks: X, A, B, and C. Each form of the survey includes the X block as well as two of the remaining blocks. A participant is then given a different form each day of the study. Designing the forms in this manner allows researchers to estimate the relationship between any chosen items, regardless of their blocks. For example, this allows the relationship between an item in block A and another in block C to be estimated and examined. \par

An important consideration here is question distribution. The X-block, being included in every form, should contain either the variables of greatest interest or those with the most predictive ability.
Placing central items in the X-block ensures they are always collected and minimizes the efficiency loss of our analyses \citep{thomas2006evaluation}. On the other hand, placing the most predictive variables in the X-block allows Multiple Imputation to more effectively impute missing values from the other blocks \citep{gottschall2012comparison}. Similarly, it is good practice to place variables which are predictive of each other in separate blocks \citep{gottschall2012comparison}. While this approach provides researchers more freedom when forming the surveys, the puzzle of question distribution introduces new concerns. When the question groupings are not obvious, pilot studies may be necessary to determine the predictive potential of and relationships between variables. The cost and burden associated with an item must also be considered when deciding which block to place it in. For example, if a critical item is overly expensive or difficult to obtain we may be better off excluding it from the X-block \citep{little2013planned}. \par

In addition, the number of blocks must be determined. Including more blocks in the design allows further increased missingness \citep{graham2006planned}. While increasing missingness may allow for further reductions in burden and cost, researchers must consider the disadvantages of more blocks. As more blocks are added, the number of possible forms increases. For example, consider two research teams. The first team divides their survey into an X-block and three other blocks of equal size, each containing 25\% of the survey. The second uses an X-block and nine others with 10\% of the items each. If the teams decide that each form will contain the X-block and two of the other blocks, then the second team will have 36 \textbf{(28? 8 choose 2?)} total forms while the first team can proceed with 3. \textbf{This is all assuming that the blocks represent the optimal question grouping. If this is not known, to account for possible relationships and correlation between questions, the teams would need to consider all possible blocks and create forms for every block configuration. For a 7 question survey, with any one question X block and the other 6 split into three blocks, there are 75 possible block configurations. For each configuration, there are 3 possible forms. So, a team would need to consider 225 different forms. Instead of creating 225 distinct forms, a research team can follow a alternative Split Form method.} \par

\textbf{This method, hereafter referred to as Split Form planned missingness and illustrated in \ref{fig:formillustration}, involves meaningfully selecting an X block of $n$ questions and then randomly excluding $k$ questions on each day. In a way treating each item as its own block \citep{silvia2014planned}. This is simple in theory, but for traditional paper surveys this is very difficult to execute. Fortunately, recent technological advances have made the use of electronic questionaires more readily accessible. In this environment form randomization can be automated and the creation of numerous unique forms is no longer an issue.} Using this method, researchers reap the benefits of including all possible variable relationships through increased forms without the difficulty caused by manually building these forms. \par

\subsection{Wave Missingness Design}
\label{sec:1.2}
While Split Form missingness can be applied to many experimental designs, the Wave Missingness design \citep{little2013planned} is unique to studies with multiple measurement occasions, or "waves". Unlike the Split Form design, where participants miss a subset of items in a single measurement occasion, the Wave Design entails randomizing participants to miss a measurement occasion entirely. An example of a Wave Missing design is shown in Figure \ref{fig:formillustration} where data is collected monthly over the course of 5 months. However, each participant is only measured at three of these occasions. \par

@@

If we have prior knowledge concerning the relationship between unobserved values, indicated by $R$, and the observed values, then Multiple Imputation can yield efficient estimates even under MNAR \citep{harel2007multiple}. However, because this is not typically the case, MI is most often implemented when data is MCAR or MAR \citep{little2014statistical}. \par

As a consequence of allowing researchers to decide which values in a dataset are observed and unobserved, PM designs grant control over the missing data mechanism. Because of this, MI is an appealing option for researchers utilizing a PM design. If implemented properly, data should conform to either the MCAR or MAR assumption. For instance, if medical researchers decide not to collect blood pressure on a completely random subset of patients there is no reason to believe the missing blood pressures differ significantly from the observed ones. This fact allows for the use of MI and other modern missing data techniques such as the Full Information Maximum Likelihood approach \citep{dempster1977maximum} \par

\subsubsection{The Multiple Imputation Procedure}
\label{sec:1.3.2}
Multiple Imputation is used to obtain statistically valid inferences in spite of missing data. MI works by replacing unobserved values with a set of plausible values in a way which allows us to account for uncertainty in the imputation process. Repeatedly using an imputation technique, every missing value in the incomplete dataset is replaced with a fixed number of imputed values. This process allows for the creation of multiple complete datasets, on which users can run complete-data procedures. If, for instance, we imputed $M$ possible values for every missing value we would obtain $M$ different completed datasets. Once created, the analysis of interest can be performed on each of them and, using rules defined by \citet{little2014statistical} and \citet{rubin2004multiple}, combine the results to obtain a single inference. \par

Suppose we are interested in estimating a certain parameter such as the mean of a variable or a regression coefficient. If $\theta$ is the parameter of interest, let $\widehat{\theta_1}, \widehat{\theta_2}, ... \widehat{\theta_M}$ be the estimates of $\theta$ and let $S_1, S_2, ... S_M$ be the corresponding estimated standard errors from the $M$ imputed datasets. We can aggregate these results using the combining rules specified by \citet{little2014statistical} and \citet{rubin2004multiple}. \par
The overall MI estimate of $\theta$ is defined as,

\begin{equation}
\widehat{\theta} = \sum_{n=1}^{M} \widehat{\theta_n}.
\end{equation}

From here we can compute the sample variance of this estimate by calculating the between-imputation component, $B_M = \frac{1}{(M-1)}\sum_{n=1}^{M} (\widehat{\theta_n} - \widehat{\theta})^2$, and the within-imputation component, $W_M = \frac{1}{(M)}\sum_{n=1}^{M} S_n^2$. These components are combined to obtain the overall sample variance of our MI estimate as follows:

\begin{equation}
T_M = W_M + \frac{M+1}{(M)}B_M.
\end{equation}

@@

When using Multiple Imputation, researchers must decide on the number of imputations, $M$, to use. Generally, as the level of missingness decreases the number of imputations needed does as well. If less than 20\% of the requisite data is missing, for instance, five or tien imputations will suffice \citep{raghunathan2015missing}. However, at its roots, MI is a simulation technique and thus, the greater $M$ is the more stable our estimates become \citep{harel2007inferences}. \citet{graham2007many} recommend that if computational power is not a limiting factor, then setting $M=100$ is good practice. \par

@@

\subsection{Simulations}
\label{sec:2.2}
The first step in the process of creating the simulated datasets was to obtain parameter estimates from \citet{pellowski2016alcohol} and to quantify the relationships between variables. These estimated relationships were treated as the "true" relationships and \textbf{one thousand (number or no???)} samples were drawn using these parameters. The purpose of simulating these datasets was to see how well they captured the true relationship between daily medication adherence on a certain day, whether or not a participant drank that day, and their beliefs about alcohol-medication interactivity (included as a Z-transformed test score). Specifically, \\

Let $Y_{ij} \sim binomial(n_{ij},p_{ij})$ be whether the $j$th person missed their medication dose on the $i$th day. Then the model of interest is the random intercept model with time effect described as follows:
\begin{align*}
logit(p_{ij}) &= \beta_{0j} + \beta_1DrinkYN_i + \beta_2ZAlcTox_i +\beta_3Day
\end{align*}
where $\beta_{0j} = \gamma_0 + \alpha_j$ and $\alpha_j \sim \mathcal{N}(0,\sigma_\alpha^2)$. \\ \\ 
Additionally, note: \\ \\
$p_{ij} = 1$ when participant $p$ did not adhere to the ART on day $i$, 0 otherwise. \\
$DrinkYN = 1$ if participant $p$  drank on day $i$, 0 otherwise.\\
$ZAlcTox =$ participant $p$'s Z-transformed score on an alcohol-antiretroviral interactive toxicity scale developed by Kalichman \cite{kalichman2009prevalence}. \\
$Day =$ Day $i$ of the study, 0-44. \\
$j =$ participant index, 1-60. \\

Based on the observed data, the true values of the parameters, $\gamma_0$, $\beta_1$, $\beta_2$, $\beta_3$, and $\sigma_\alpha^2$ were found to be -2.29, 0.28, 0.64, 0.01, and \textbf{1.0???} respectively. \par
Once the samples were drawn, they were transformed under the three different missingness structures. These were the Split-Form and Wave Missingness designs previously mentioned as well as a variation on the Split-Form design where a few measurement occasions (here, two days per participant) are fully observed. To clarify this difference, see the hypothetical application of these designs to a five question multi-day survey in \textbf{Table \ref{tab:table14} (do we want the illustration here instead? or in both places?)}. As in the previous tables, "1" indicates an observed occasion while "0" indicates a missing one. We see that the Split Form design entails missing two items on each of the days. The altered Split Form, while similar, is fully measured on two of the measurement occasions. Because there are days which are fully observed, the relationships between variables may be clearer, allowing the imputation procedure to run more effectively. On the other hand, the Wave Missingness design shows two days in which no questions are asked but the full questionnaire is given on the remaining days. \par

@@

Additionally, for each different Missingness pattern we imposed low, medium, and high levels of missingness roughly corresponding to 2, 4, and 6 missing questions. Previous work by \citet{rhemtulla2014planned} found that for low levels of missingness Split Form design outperforms the Wave Missingness design in capturing latent growth curve model parameters. However, because only low-levels of missingness are tested it is unknown whether this relationship holds at moderate or high levels. By including these different levels of missingness the performance at all levels of the missingness spectrum will be measured. \par
Once missingness was created, Multiple Imputation techniques were applied to estimate the relationship between medication adherence and the other variables in our model. Because the missing data mechanism is controled, the data is known to be MCAR. This fact allows for the principled use of MI techniques. Using the R package $\{$mice$\}$ \citep{mice2011imputation} the binary variables (Q1-Q7) were imputed using a logistic model based approach while the count variable (Q8) was imputed using predictive mean matching \textbf{was it??? i don't think so, it was in the X block originally but double check!}. \par

@@

\subsubsection{Increased Sample Size}
In order to evaluate the impact of sample size on estimation, the original sample size of 60 participants was expanded to 120. The rest of the simulation process was carried out as before. As noted earlier, the cost savings gleaned from the PM designs could allow us to enroll more participants. Thus, even if one design performed more optimally than the others, if all designs also perform acceptably well at high samples, researchers can design missingness in whichever method is most convenient. Furthermore, because Wave Missingness allows for greater savings, an increase in sample size is more reasonable for that type of missingness. \par
\subsubsection{Low Inter-Survey Correlation}
For our second simulation scenario we looked to determine the effect of decreased inter-survey correlation on which PM design would perform best. To clarify, consider the fact that answers given on a particular day may be correlated with one another- the two questions on housing insecurity, for example, are likely to be "yes" on the same days. While this fact was accounted for in the original simulations, it was not in this set of simulations. If the probability of answering "yes" on Question 2, for instance, was a function of the answer to Question 1 and the baseline variables in the original simulations, then in this scenario it is a function of only the baseline variables. This implies that each of the questions are conditionally independent of each other given the baseline variables. If researchers have a questionnaire where they do not expect any of the answers to be obviously related the results here may help to choose which missingness design would be most optimal. To investigate this, the correlation of the variables was scaled by 0.5.\par
\subsubsection{High Time Effect}
In the final scenario, the goal was to determine which PM design would lead to the best estimates when time has a disproportionately large effect on our outcome variable. To do this we changed the value of the parameters $\gamma_0$, $\beta_1$, $\beta_2$, and $\beta_3$ to -4.56, 0.14, 0.32, and 0.01 respectively. Note that instead of increasing the value of $\beta_3$, the slope for time effect, all the other parameters were decreased. This was done because $\beta_3$ was already fairly large and increasing its value further would cause the probability of missing a dose to rapidly approach 1. Decreasing the other variables effects ensured that time has a disproportionate effect without this problem. The results of this section may prove useful to researchers who believe that the overall passage of time plays a larger role in their outcome than a measurement taken at one specific point. \par

\section{Results}
\label{sec:3}
\subsection{Primary Simulation}
\label{sec:3.1}
The first set of simulations run contained the parameters in the random-intercept logistic model with $\gamma_0 = -2.28$, $\beta_1 = 0.28$, $\beta_2 = 0.64$, and $\beta_3 = 0.01$. The effects of the variables here are comparable to those found in the original dataset. This fact implies that we can interpret the results here to show what may have occurred if the original study had implemented a PM design. Before we imposed a missingness structure on the 1,000 simulated samples we obtained the full data results shown in Table \ref{tab:prelim} \textbf{not here yet}. Note that the full data results for the other simulation scenarios are also provided in this table for ease of reference.

Of note in the complete-data primary analysis is that the coverage of the 95\% CIs is larger than expected for the parameter estimates and the percent bias for the $\beta_1$ estimate is somewhat large. This increased bias is likely due to the difficulties inherent in estimating logistic regression models with rare event data \citep{king2001logistic}. Nevertheless, this must be kept in mind when analyzing the results from the PM designs provided in Table \ref{tab:primary}. \textbf{In Table \ref{tab:primary} as in Tables \ref{tab:sampsize}, \ref{tab:lowcorr}, and \ref{tab:hightime} bold values indicate ``unacceptable" results.} That is, bias, percent bias, or MSE more than 40\% greater than the complete-data value or a CI coverage of less than 90\%.

@@

For the Split Form and Altered Split Form designs seen in Table \ref{tab:table3}, note that the greatest loss in efficiency was on the $\beta_1$ estimate. While the CI coverage remained high at all levels of missingness for the two designs, the bias, percent bias, and MSE reached unacceptable levels at the high missingness level. We also point out that the Altered Split Form design did not perform significantly better than the regular Split Form design. While some individual measurements outperformed those from the Split Form design, bias on $\beta_1$ for example, it underperformed in others, such as percent bias on $\beta_3$. In reference to the FMI on the $\beta_1$ estimate, we can see that at low and medium missingness both Split Form designs performed well. That is, there was less variance in the results from th imputed datasets than we expected given the amount of missing data.\par

>>>>

@@

% Predefine some figures
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}

{\ttfamily\noindent\color{warningcolor}{\#\# Warning: package 'dplyr' was built under R version 3.4.3}}

{\ttfamily\noindent\color{warningcolor}{\#\# Warning: package 'kableExtra' was built under R version 3.4.4}}\end{kframe}
\end{knitrout}

@@

\author{Josephine Wood \and Gregory J. Matthews \and Jennifer Pellowski \and Ofer Harel}

\institute{J. Wood \and G. J. Matthews
          \at Loyola University Chicago \\
          1032 W. Sheridan Rd.\\
          Chicago IL, 60660 USA\\
          Tel: (773)508-3558
          \and
          O. Harel 
				\at University of Connecticut \\
						215 Glennbrook Rd. Unit 420\\
						Storrs CT, 06269-4120 USA
              Tel: (860)486-6989 \\
		      \email{ofer.harel@uconn.edu}  
					\and
					J. Pellowski
					\at International Health Institute\\ 
					Brown School of Public Health\\
					Department of Psychiatry and Human Behavior\\
					The Warren Alpert Medical School of Brown University.  
}

@@

\section{Introduction}
\label{intro}

Often, studies are conducted longitudinally; measuring participants on numerous separate occasions. This allows researchers to observe how a measurement may change over time. In theory, administering a survey once a month would be successful. However, in practice, as surveys become longer and more frequent, participants may begin to provide ingenuine responses or drop out of the study entirely. To combat this issue, researchers may attempt to shorten the survey or decrease the measurement frequency. Merely truncating a survey results in lost information and infrequent administration does not provide a clear picture of participants' evolution over time. In answer to a difficult question, many researchers have turned to planned missingness when designing experiments.

Planned missingness (PM) designs are a type of experimental design in which some of the data is predetermined to be missing. While it may seem counterproductive to deliberately omit data from the collection process, PM designs have been shown to reduce both participant burden \citep{graham2006planned, saris20048} and the costs of a study \citep{graham2001planned}. Moreover, PM designs can actually increase the quality of collected data by increasing the breadth of a study \citep{little2013planned} as well as decreasing the effects of assessment reactivity and rates of unplanned missingness \citep{harel2015designed}. We further note that while PM designs decrease the efficiency of estimators \citep{rhemtulla2016asymptotic}, decreasing the cost per participant may allow researchers to enroll more participants in a study, offsetting some of the efficiency loss. \par

There have been many different incarnations of the PM design, each crafted to best suit the study at hand. For example, PM designs have been used in the study of longitudinal growth curves \citep{graham2001planned, mcardle1997expanding}, business research \citep{shoemaker1973principles}, behavioral studies \citep{harel2015designed}, biostatistics \citep{andres2006partial, wacholder1994partial}, and educational assessments \citep{zeger1997efficient, sirotnik1977incidence}. This paper focuses on two of the simplest and most applicable PM designs; the Split Form design \citep{raghunathan1995split} and the Wave Missingness design \citep{little2013planned} and their application in a longitudinal study utilizing a multi-level regression model, as it lacks coverage in recent literature. While this paper covers these designs in more detail later, broadly speaking, the Split Form design entails measuring fewer points on a single measurement occasion \citep{raghunathan1995split}. The Wave Missingness design, on the other hand, requires missing an entire measurement occasion within a longitudinal study \citep{graham2001planned}.

@@

Split Form missingness designs \citep{raghunathan1995split}, also  known as Multi-Form \citep{little2013planned} or Matrix Sampling designs \citep{thomas2006evaluation}, are aptly named. Instead of giving participants one large survey, researchers create multiple forms, each containing a subset of the full survey's questions. How these subsets are chosen varies. A common form setup is three-form Block Design \citep{graham1996maximizing} in which the full survey is split into four blocks: X, A, B, and C. Each form of the survey includes the X block as well as two of the remaining blocks. A participant is then given a different form each day of the study. Designing the forms in this manner allows researchers to estimate the relationship between any chosen items, regardless of their blocks. For example, this allows the relationship between an item in block A and another in block C to be estimated and examined. \par

An important consideration here is question distribution. The X-block, being included in every form, should contain either the variables of greatest interest or those with the most predictive ability.

Placing central items in the X-block ensures they are always collected and minimizes the efficiency loss of our analyses \citep{thomas2006evaluation}. On the other hand, placing the most predictive variables in the X-block allows Multiple Imputation to more effectively impute missing values from the other blocks \citep{gottschall2012comparison}. Similarly, it is good practice to place variables which are predictive of each other in separate blocks \citep{gottschall2012comparison}. While this approach provides researchers more freedom when forming the surveys, the puzzle of question distribution introduces new concerns. When the question groupings are not obvious, pilot studies may be necessary to determine the predictive potential of and relationships between variables. The cost and burden associated with an item must also be considered when deciding which block to place it in. For example, if a critical item is overly expensive or difficult to obtain we may be better off excluding it from the X-block \citep{little2013planned}. \par

In addition, the number of blocks must be determined. Including more blocks in the design allows further increased missingness \citep{graham2006planned}. While increasing missingness may allow for further reductions in burden and cost, researchers must consider the disadvantages of more blocks. As more blocks are added, the number of possible forms increases. For example, consider two research teams. The first team divides their survey into an X-block and three other blocks of equal size, each containing 25\% of the survey. The second uses an X-block and nine others with 10\% of the items each. If the teams decide that each form will contain the X-block and two of the other blocks, then the second team will have 36 total forms while the first team can proceed with 3. 
%\textbf{This is all assuming that the blocks represent the optimal question grouping. If this is not known, to account for possible relationships and correlation between questions, the teams would need to consider all possible blocks and create forms for every block configuration. For a 7 question survey, with any one question X block and the other 6 split into three blocks, there are 75 possible block configurations. For each configuration, there are 3 possible forms. So, a team would need to consider 225 different forms. Instead of creating 225 distinct forms, a research team can follow a alternative Split Form method.} \par

This is simple in theory, but for traditional paper surveys this is very difficult to execute. Fortunately, recent technological advances have made the use of electronic questionaires more readily accessible. In this environment form randomization can be automated and the creation of numerous unique forms is no longer an issue. This ability allows another method, hereafter referred to as Split Form planned missingness and illustrated in Figure \ref{fig:formillustration}, to be implemented. Split Form is the random exclusion of $k$ questions on a given day, treating each item as its own block \citep{silvia2014planned}.  Using this method, researchers reap the benefits of including all possible variable relationships through increased forms without the difficulty caused by manually building these forms. \par

\subsection{Wave Missingness Design}
\label{sec:1.2}
While Split Form missingness can be applied to many experimental designs, the Wave Missingness design \citep{little2013planned} is unique to studies with multiple measurement occasions, or ``waves". Unlike the Split Form design, where participants miss a subset of items in a single measurement occasion, the Wave Design entails randomizing participants to miss a measurement occasion entirely. An example of a Wave Missing design is shown in Figure \ref{fig:formillustration} where data is collected monthly over the course of 5 months. However, each participant is only measured at three of these occasions. \par

@@

If we have prior knowledge concerning the relationship between unobserved values, indicated by $R$, and the observed values, then Multiple Imputation can yield efficient estimates even under MNAR \citep{harel2007multiple}. However, because this is not typically the case, MI is most often implemented under MCAR or MAR assumptions \citep{little2014statistical}. \par

As a consequence of allowing researchers to decide which values in a dataset are observed and unobserved, PM designs grant control over the missing data mechanism. Because of this, MI is an appealing option for researchers utilizing a PM design. If implemented properly, data should conform to either the MCAR or MAR assumption. For instance, if medical researchers decide not to collect blood pressure on a completely random subset of patients there is no reason to believe the missing blood pressures differ significantly from the observed ones. This fact allows for the use of MI and other modern missing data techniques such as the Full Information Maximum Likelihood approach \citep{dempster1977maximum}. \par

\subsubsection{The Multiple Imputation Procedure}
\label{sec:1.3.2}
Multiple Imputation is used to obtain statistically valid inferences in spite of missing data. MI works by replacing unobserved values with a set of plausible values in a way which allows us to account for uncertainty in the imputation process. Repeatedly using an imputation technique, every missing value in the incomplete dataset is replaced with a fixed number of imputed values. This process allows for the creation of multiple complete datasets, on which users can run complete-data procedures. If, for instance, we imputed $M$ possible values for every missing value we would obtain $M$ different completed datasets. Once created, the analysis of interest can be performed on each of them and, using Rubin's rules detailed in \citet{rubin2004multiple}, combine the results to obtain a single inference. \par

Suppose we are interested in estimating a certain parameter such as the mean of a variable or a regression coefficient. If $\theta$ is the parameter of interest, let $\widehat{\theta_1}, \widehat{\theta_2}, ... \widehat{\theta_M}$ be the estimates of $\theta$ and let $S_1, S_2, ... S_M$ be the corresponding estimated standard errors from the $M$ imputed datasets. We can aggregate these results using the combining rules specified in \citet{rubin2004multiple}. \par
The overall MI estimate of $\theta$ is defined as,

\begin{equation}
\widehat{\theta} = \sum_{n=1}^{M} \frac{1}{M} \widehat{\theta_n}.
\end{equation}

From here we can compute the sample variance of this estimate by calculating the between-imputation component, $B_M = \frac{1}{(M-1)}\sum_{n=1}^{M} (\widehat{\theta_n} - \widehat{\theta})^2$, and the within-imputation component, $W_M = \frac{1}{M}\sum_{n=1}^{M} S_n^2$. These components are combined to obtain the overall sample variance of our MI estimate as follows:

\begin{equation}
T_M = W_M + \frac{M+1}{M}B_M.
\end{equation}

@@

When using Multiple Imputation, researchers must decide on the number of imputations, $M$, to use. Generally, as the level of missingness decreases the number of imputations needed does as well. \textbf{Greg: IS THIS TRUE?} If less than 20\% of the requisite data are missing, for instance, five or ten imputations will suffice \citep{raghunathan2015missing}. However, at its roots, MI is a simulation technique and thus, the greater $M$ is the more stable our estimates become \citep{harel2007inferences}. \citet{graham2007many} recommend that if computational power is not a limiting factor, then setting $M=100$ is good practice. \par

@@

\subsection{Simulations}
\label{sec:2.2}
The first step in the process of creating the simulated datasets was to obtain parameter estimates from \citet{pellowski2016alcohol} and to quantify the relationships between variables. These estimated relationships were treated as the "true" relationships and one thousand samples were drawn using these parameters. The purpose of simulating these datasets was to see how well they captured the true relationship between daily medication adherence on a certain day, whether or not a participant drank that day, and their beliefs about alcohol-medication interactivity (included as a Z-transformed test score). Specifically, \\

Let $Y_{ij} \sim binomial(n_{ij},p_{ij})$ be whether the $j$th person missed their medication dose on the $i$th day. Then the model of interest is the random intercept model with time effect described as follows:
\begin{align*}
logit(p_{ij}) &= \beta_{0j} + \beta_1DrinkYN_i + \beta_2ZAlcTox_i +\beta_3Day
\end{align*}
where $\beta_{0j} = \gamma_0 + \alpha_j$ and $\alpha_j \sim N(0,\sigma_\alpha^2)$. \\ \\ 
Additionally, note: \\ \\
\begin{itemize}
\item $p_{ij} = 1$ when participant $p$ did not adhere to the ART on day $i$, 0 otherwise \\
\item $DrinkYN = 1$ if participant $p$  drank on day $i$, 0 otherwise\\
\item $ZAlcTox =$ participant $p$'s Z-transformed score on an alcohol-antiretroviral interactive toxicity scale developed by Kalichman \cite{kalichman2009prevalence} \\
\item $Day =$ Day $i$ of the study, 0-44 \\
\item $j =$ participant index, 1-60 \\
\end{itemize}.

Based on the observed data, the true values of the parameters, $\gamma_0$, $\beta_1$, $\beta_2$, $\beta_3$, and $\sigma_\alpha^2$ were found to be -2.29, 0.28, 0.64, 0.01, and 1.12, respectively. \par
Once the samples were drawn, missingness was added under three different structures. These were the Split-Form and Wave Missingness designs previously mentioned as well as a variation on the Split-Form design where a few measurement occasions (here, two days per participant) are fully observed. To clarify this difference, see the hypothetical application of these designs to a five question multi-day survey in Table \ref{tab:table14} \textbf{Greg: I think we need either only the table or only the illustration.  Not sure we need both.}. As in the previous tables, "1" indicates an observed occasion while "0" indicates a missing one. We see that the Split Form design entails missing two items on each of the days. The altered Split Form, while similar, is fully measured on two of the measurement occasions. Because there are days which are fully observed, the relationships between variables may be clearer, allowing the imputation procedure to run more effectively. On the other hand, the Wave Missingness design shows two days in which no questions are asked but the full questionnaire is given on the remaining days. \par

@@

%FIX THIS: THIS SOUNDS LIKE WE ADDED MORE MISSINGNESS. FIXED
For each different designed missingness pattern we imposed low, medium, and high levels of missingness roughly corresponding to 2, 4, and 6 missing questions. Previous work by \citet{rhemtulla2014planned} found that for low levels of missingness Split Form design outperforms the Wave Missingness design in capturing latent growth curve model parameters. However, because only low-levels of missingness are tested it is unknown whether this relationship holds at moderate or high levels. By including these different levels of missingness the performance at all levels of the missingness spectrum will be measured. \par
Once missingness was created, Multiple Imputation techniques were applied to estimate the relationship between medication adherence and the other variables in our model. Because the missing data mechanism is controled, the data is known to be MCAR. This fact allows for the principled use of MI techniques. Using the R package \texttt{mice} \citep{mice2011imputation} the binary variables (Q1-Q7) were imputed using a logistic model based approach while the count variable (Q8) was imputed using predictive mean matching %\textbf{was it??? i don't think so, it was in the X block originally but double check!Greg: Yes it was}. 
\par

@@

\subsubsection{Increased Sample Size}
In order to evaluate the impact of sample size on estimation, the original sample size of 60 participants was expanded to 120. The rest of the simulation process was carried out as before. As noted earlier, the cost savings gleaned from the PM designs could allow us to enroll more participants. Thus, even if one design performed more optimally than the others, if all designs also perform acceptably well at high samples, researchers can design missingness in whichever method is most convenient. Furthermore, because Wave Missingness allows for greater savings, an increase in sample size is more reasonable for that type of missingness. \par
\subsubsection{Low Inter-Survey Correlation}
For our second simulation scenario we looked to determine the effect of decreased inter-survey correlation on which PM design would perform best. To clarify, consider the fact that answers given on a particular day may be correlated with one another- the two questions on housing insecurity, for example, are likely to be "yes" on the same days. While this fact was accounted for in the original simulations, it was not in this set of simulations. If the probability of answering "yes" on Question 2, for instance, was a function of the answer to Question 1 and the baseline variables in the original simulations, then in this scenario it is a function of only the baseline variables. This implies that each of the questions are conditionally independent of each other given the baseline variables. If researchers have a questionnaire where they do not expect any of the answers to be obviously related the results here may help to choose which missingness design would be most optimal. To investigate this, the correlation of the variables was scaled by 0.5.\par
\subsubsection{High Time Effect}
In the final scenario, the goal was to determine which PM design would lead to the best estimates when time has a disproportionately large effect on our outcome variable. To do this held the values of $\gamma_0$, $\beta_1$, $\beta_2$ the same and increaed the effect of time (i.e. $\beta_3$) by ten times.  The results of this section may prove useful to researchers who believe that the overall passage of time plays a larger role in their outcome than a measurement taken at one specific point. \par

\section{Results}
\label{sec:3}
\subsection{Primary Simulation}
\label{sec:3.1}
The first set of simulations run contained the parameters in the random-intercept logistic model with $\gamma_0 = -2.28$, $\beta_1 = 0.28$, $\beta_2 = 0.64$, and $\beta_3 = 0.01$. The effects of the variables here are comparable to those found in the original dataset. This fact implies that we can interpret the results here to show what may have occurred if the original study had implemented a PM design. Before we imposed a missingness structure on the 1,000 simulated samples we obtained the full data results shown in Table \ref{tab:prelim}. Note that the full data results for the other simulation scenarios are also provided in this table for ease of reference.

@@

Of note in the complete-data primary analysis is that the coverage of the 95\% CIs is larger than expected for the parameter estimates and the percent bias for the $\beta_1$ estimate is somewhat large. This increased bias is likely due to the difficulties inherent in estimating logistic regression models with rare event data \citep{king2001logistic}. Nevertheless, this must be kept in mind when analyzing the results from the PM designs provided in Table \ref{tab:primary}. %\textbf{In Table \ref{tab:primary} as in Tables \ref{tab:sampsize}, \ref{tab:lowcorr}, and \ref{tab:hightime} bold values indicate ``unacceptable" results.} That is, bias, percent bias, or MSE more than 40\% greater than the complete-data value or a CI coverage of less than 90\%.

\caption{\label{tab:primary}\label{} Primary Results (N = 60)}
\centering
\resizebox{\linewidth}{!}{\begin{tabular}[t]{l|l|l|r|l|r|l|r|r}
\hline
Design & Level & Parameter & Bias & Pct Bias & MSE & Coverage & CI Length & FMI\\
\hline
 &  & $\gamma_0$ & 0.042 & -1.82\% & 0.018 & 98.8\% & 0.702 & 0.002\\

 &  & $\beta_1$ & -0.071 & -25.12\% & 0.269 & 98.3\% & 2.534 & 0.236\\

 &  & $\beta_2$ & 0.008 & 1.25\% & 0.024 & 93.2\% & 0.585 & 0.000\\

 & \multirow{-4}{*}{\raggedright\arraybackslash Low} & $\beta_3$ & 0.000 & -0.89\% & 0.000 & 94.4\% & 0.017 & 0.001\\

 &  & $\gamma_0$ & 0.039 & -1.68\% & 0.018 & 99\% & 0.709 & 0.011\\

 &  & $\beta_1$ & -0.102 & -36.19\% & 0.271 & 99.8\% & 3.500 & 0.534\\

 &  & $\beta_2$ & 0.008 & 1.21\% & 0.024 & 93.3\% & 0.586 & 0.002\\

 & \multirow{-4}{*}{\raggedright\arraybackslash Medium} & $\beta_3$ & 0.000 & -0.42\% & 0.000 & 94.6\% & 0.017 & 0.007\\

 &  & $\gamma_0$ & 0.028 & -1.24\% & 0.027 & 98.5\% & 0.791 & 0.030\\

 &  & $\beta_1$ & -0.107 & -38.05\% & 0.418 & 98.8\% & 4.268 & 0.630\\

 &  & $\beta_2$ & 0.012 & 1.87\% & 0.027 & 93.2\% & 0.613 & 0.005\\

\multirow{-12}{*}{\raggedright\arraybackslash Split Form} & \multirow{-4}{*}{\raggedright\arraybackslash High} & $\beta_3$ & 0.000 & -0.47\% & 0.000 & 94.6\% & 0.020 & 0.020\\
\cline{1-9}
 &  & $\gamma_0$ & 0.042 & -1.83\% & 0.018 & 99\% & 0.701 & 0.001\\

 &  & $\beta_1$ & -0.069 & -24.4\% & 0.283 & 98.3\% & 2.508 & 0.218\\

 &  & $\beta_2$ & 0.008 & 1.27\% & 0.024 & 93.3\% & 0.585 & 0.000\\

 & \multirow{-4}{*}{\raggedright\arraybackslash Low} & $\beta_3$ & 0.000 & -0.95\% & 0.000 & 94.4\% & 0.017 & 0.001\\
